\textbf{Proof Reconstruction in Classical Propositional Logic}

### 2017-04-24 slides/ideas.tex
[ slides ] diagrams added.

\documentclass[hyperref={pdfpagelabels=false}]{beamer}
\beamertemplatenavigationsymbolsempty
\usepackage[utf8x]{inputenc}
\usepackage{amsmath, amsthm, amssymb}
%\usepackage{bm}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{lmodern}
\usepackage{url}
\usepackage[T1]{fontenc}

\definecolor{aliceblue}{rgb}{0.94, 0.97, 1.0}
\definecolor{bondiblue}{rgb}{0.0, 0.58, 0.71}
\definecolor{cadet}{rgb}{0.33, 0.41, 0.47}

\lstset{
    aboveskip=10pt
  , backgroundcolor=\color{aliceblue}
  , basicstyle=\ttfamily
  , belowskip=9pt
  , frame=trbl
  , keywords=[2]{axiom, conjecture,inference, negated_conjecture, plain}
  , keywords=[3]{fof, cnf}
  , keywords=[4]{canonicalize, resolve, conjunct, strip, simplify, negate, clausify}
  , keywordstyle=[2]\color{cadet}%\underbar,%\bfseries
  , keywordstyle=[3]\color{bondiblue}\bfseries
  , keywordstyle=[4]\bfseries
  , showstringspaces=false
}

\newcommand{\problemtptp}[3][c]{\lstinputlisting[caption=#2, firstline=#3, label=custom#1]{#2}}
\newcommand{\solutiontstp}[2][c]{\lstinputlisting[caption=#2, firstline=5, label=custom#1]{#2}}

\title{Proof Reconstruction with Athena\\(Work in Progress)}

\date{Agda Implementors’ Meeting XXV}

\author{Jonathan Prieto-Cubides}

\institute{
Advisor: Andr\'es Sicard-Ram\'irez\\[3mm]
EAFIT University\\
Medell\'in, Colombia}

\begin{document}

\setcounter{page}{1}
\maketitle

% Context
% - SledgeHammer
% - Metis
% - Problem with Wallmeinster finding the code.
% - equinox (intuisionistic version)

\begin{frame}{Keywords}
  \begin{block}{Proof Assistant}
    is a software program that work together with a human in order to give formal proofs to problems in a wide range of topics.\\
    - Agda, Isabelle/HOL, HOL Light, Coq.
  \end{block}
  \begin{block}{\text{A}utomatic \textbf{T}heorem \textbf{P}rover}
    Is software program that tries to prove conjecture based on a list of hypothesis.\\
    - Metis, Z3, Vampire, EProver, iProver, SPSS.
  \end{block}
\end{frame}

\begin{frame}{Athena}
  Is a Haskell program that translates proofs given by Metis ATP in TSTP format
  to Agda code.

  \begin{figure}
    \includegraphics[scale=0.85]{diagram.pdf}
    \label{im:athena}
    \caption{Proof reconstruction work flow.}
  \end{figure}
\end{frame}

@@

\begin{frame}{References}

\end{document}

>>>>

\begin{frame}{Keywords}

 \tikz[overlay,remember picture]
  \node[opacity=0.2]at (0.8\textwidth, 0.5cm)
    {\includegraphics[width=0.3\textwidth]{figures/metis}};

  \begin{block}{Proof Assistant}
    Is a software program that work together with a human in order to give formal proofs to problems in a wide range of topics.\\
    - Agda, Isabelle/HOL, HOL Light, Coq.
  \end{block}
  \begin{block}{\text{A}utomatic \textbf{T}heorem \textbf{P}rover}
    Is software program that tries to prove conjecture based on a list of hypothesis.\\
    - Metis, Z3, Vampire, EProver, iProver, SPSS.
  \end{block}
\end{frame}

\begin{frame}{Athena}{asdf}
  Is a Haskell program that translates proofs given by Metis ATP in TSTP format
  to Agda code.

  \begin{figure}
    \includegraphics[scale=0.85]{figures/diagram.pdf}
    \label{im:athena}
    \caption{Proof reconstruction work flow.}
  \end{figure}
\end{frame}

@@

### 2017-04-24 slides/slides.tex
[ slides ] diagrams added.

% --------------------------------------------------------------------------
% My palette.
% --------------------------------------------------------------------------
\definecolor{energy}{RGB}{49,247,250}
\definecolor{delicate}{RGB}{67,179,223}
\definecolor{faded}{RGB}{76,117,195}
\definecolor{plum}{RGB}{87,78,164}
\definecolor{petunias}{RGB}{109,80,139}
% --------------------------------------------------------------------------

@@

\setbeamercolor{institute}{fg=blu}
% \setbeamercolor{subtitle}{fg=blu}
% \setbeamercolor{titlelike}{fg=blu}
% \setbeamerfont{footnote}{size=\tiny}
% \setbeamercolor{block title}{bg=blue,fg=blu}
% \setbeamercolor{block body}{bg=aliceblue}
% \setbeamercolor{item}{fg=blu} % color of bullets
% \setbeamercolor{subitem}{fg=blu}
% \setbeamercolor{itemize/enumerate subbody}{fg=blu}
% \setbeamertemplate{itemize subitem}{{\textendash}}
% \setbeamerfont{itemize/enumerate subbody}{size=\footnotesize}
% \setbeamerfont{itemize/enumerate subitem}{size=\footnotesize}

@@

\usepackage[english]{babel}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{bussproofs}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage{lstagda}
\usepackage{tikz}
\usepackage{url}

@@

% --------------------------------------------------------------------------
% Title Page
\title{\textbf{Proof Reconstruction in Classical Propositional Logic}}
\date{Agda Implementors’ Meeting XXV\\
May 9-15th
}
\author{Jonathan Prieto-Cubides}

@@

@@

\begin{frame}[plain]
  \titlepage
\end{frame}

@@

\begin{frame}
Proof Term Reconstruction

At the moment, the communication between Agda and
the ATPs is unidirectional because we use the ATPs as oracles

@@

\begin{frame}[fragile]{TPTP Syntax}{Thousands of Problems for Theorem Provers}

 \tikz[overlay,remember picture]
  \node at (0.92\textwidth, 0cm)
    {\includegraphics[width=0.15\textwidth]{figures/tptp}};

@@

\begin{frame}[fragile]{TPTP Examples}{Prop-Pack Repository}
Problems in Propositional Logic:
  \begin{itemize}
    \item $p ⊢ p$
\begin{tptp}
$ cat basic-4.tptp
fof(a, axiom, p).
fof(goal, conjecture, p).
\end{tptp}
    \item $p ∧ q ⊢ q ∧ p$
\begin{tptp}
$ cat conj-3.tptp
fof(a, axiom, p & q).
fof(goal, conjecture, q & p).
\end{tptp}
    \item $⊢ ¬ (p ∧ ¬ p) ∨ (q ∧ ¬ q)$
\begin{tptp}
$ cat neg-7.tptp
fof(goal, conjecture, ~ ((p & ~ p) | (q & ~ q))).
\end{tptp}
    \end{itemize}
\end{frame}

@@

\begin{frame}[fragile]{TSTP Syntax}
 \tikz[overlay,remember picture]
  \node at (0.92\textwidth, -0.5cm)
    {\includegraphics[width=0.15\textwidth]{figures/tstp}};

A TSTP derivation\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/QuickGuide/Derivations.html}}
\begin{itemize}
  \item Is a Directed Acyclic Graph where
  \begin{itemize}
    \item[\texttt{leaf}] is a formulae from the TPTP input
    \item[\texttt{node}] is a formulae inferred from parent formulae
    \item[\texttt{root}] the final derived formulae
  \end{itemize}
  \item Is a list of annotated formulae:
  {\footnotesize
    \begin{center}
\begin{verbatim}
language(name,role,formula,source[,useful info]).
\end{verbatim}
  \end{center}}
\begin{itemize}
  \item[\texttt{source}] typically is a inference record:
  {\footnotesize \begin{verbatim}
  inference(rule, [], [])
  \end{verbatim}}
  % Where each slot corresponds to
\begin{itemize}
    \item[-] The inference rule name,
    \item[-] a list of useful inference information, and
    \item[-] a list of references to its parent formulae.
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{TSTP Example}
 \tikz[overlay,remember picture]
  \node at (0.93\textwidth, -0.3cm)
    {\includegraphics[width=0.15\textwidth]{figures/metis}};

@@

\begin{frame}[fragile]{DAG for the previous TSTP derivation found by \texttt{Metis} ATP}
\vfill
\begin{center}
\includegraphics[height=0.8\textheight]{figures/derivation.png}
\end{center}
\end{frame}

>>>>

% --------------------------------------------------------------------------
% My palette.
% --------------------------------------------------------------------------
\definecolor{energy}{RGB}{49,247,250}
\definecolor{delicate}{RGB}{67,179,223}
\definecolor{faded}{RGB}{76,117,195}
\definecolor{plum}{RGB}{87,78,164}
\definecolor{petunias}{RGB}{109,80,139}
\definecolor{letour}{RGB}{101,41,105}

@@

% \setbeamercolor{institute}{fg=blu}
% \setbeamercolor{subtitle}{fg=blu}
% \setbeamercolor{titlelike}{fg=blu}
\setbeamerfont{footnote}{size=\tiny}
\setbeamercolor{footnote}{fg=gray}
% \setbeamercolor{block title}{bg=blue,fg=blu}
% \setbeamercolor{block body}{bg=aliceblue}
% \setbeamercolor{item}{fg=blu} % color of bullets
% \setbeamercolor{subitem}{fg=blu}
% \setbeamercolor{itemize/enumerate subbody}{fg=blu}
% \setbeamertemplate{itemize subitem}{{\textendash}}
% \setbeamerfont{itemize/enumerate subbody}{size=\footnotesize}
% \setbeamerfont{itemize/enumerate subitem}{size=\footnotesize}

@@

\usepackage[english]{babel}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{bussproofs}
\usepackage{fancyvrb}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage{lstagda}
\usepackage{tikz}
\usepackage{url}

@@

% --------------------------------------------------------------------------
% References
\usepackage[autostyle]{csquotes}
\usepackage[
    backend=biber
  , style=authoryear-icomp
  , sortlocale=en_US
  , natbib=true
  , url=false
  , doi=true
  , eprint=false
]{biblatex}
\addbibresource{ref.bib}
\renewcommand*{\nameyeardelim}{\addcomma\addspace}
\usepackage{silence}
\WarningFilter{biblatex}{Patching footnotes failed}

% --------------------------------------------------------------------------
% Title Page
\title{\textbf{Proof Reconstruction in Classical Propositional Logic}}
\date{Agda Implementors’ Meeting XXV\\
May 9-15th}
\author{Jonathan Prieto-Cubides}

@@

\newsavebox\agdapragma

@@

\begin{frame}[plain]
% \tikz[overlay,remember picture] \node[opacity=0.05, at=(6,6)] {
% $Γ ⊢ ϕ$
% };
\titlepage
\end{frame}

@@

\begin{lrbox}{\agdapragma}
\begin{lstlisting}
$ cat Or.agda
module Or where

data _or_ (A B : Set) : Set where
  inj1 : A -> A or B
  inj2 : B -> A or B

postulate
  A B    : Set
  or-comm : A or B -> B or A
{-# ATP prove or-comm #-}
\end{lstlisting}
\end{lrbox}

\begin{frame}[fragile]{Translation}

\only<1-2>{At the moment, the communication between Agda and
the ATPs is unidirectional because the ATPs are being used as oracles.\citep{Sicard2015}.}
\only<1-2>{\vfill}

\begin{tikzpicture}[scale=2]
\only<1->{ \node(agda-pragmas) at (0,0){\includegraphics{figures/agda-pragmas.pdf}}};
\only<1>{
  \node[fill=aliceblue, rectangle](atp-pragma) at (3,0){
  \scalebox{0.7}{\usebox\agdapragma}
  }};

\only<2->{ \node(eagda) at (2,0){
  \includegraphics{figures/eagda.pdf}
  \footnote{Development version of Agda in order to handle a new built-in ATP-pragma.
    \url{https://github.com/asr/eagda}}
  }};

\only<3->{ \node(apia) at (4,0){
  \includegraphics{figures/apia.pdf}
  \footnote{Haskell program for proving first-order theorems written in Agda using ATPs.
    \url{https://github.com/asr/apia}}}
  };

\only<3->{ \node(tptp) at (2,-2){\includegraphics{figures/tptp.pdf}}};

\only<3->{ \node(atp) at (4,-2){\includegraphics{figures/atp.pdf}}};

\only<2->{\draw[->,ultra thick] (agda-pragmas) to (eagda)};

\only<3->{\draw[->,ultra thick] (eagda) to (apia)};
\only<3->{\draw[->, very thick, dashed, gray] (apia) to (tptp)};
\only<3->{\draw[->, very thick, dashed, gray] (tptp) to (atp)};
\only<3->{\draw[->, very thick, dashed, gray] (atp) to (apia)};

\end{tikzpicture}

@@

\begin{frame}[fragile]{TPTP Syntax}{Thousands of Problems for Theorem Provers}
\label{tptp-syntax}
 \tikz[overlay,remember picture]
  \node at (0.92\textwidth, 0cm)
    {\includegraphics[width=0.15\textwidth]{figures/tptp}};

@@

\begin{frame}[fragile]{TPTP Examples}{Prop-Pack Repository}
\label{tptp-examples}
Problems in Propositional Logic:
  \begin{itemize}
    \item $p ⊢ p$
\begin{tptp}
$ cat basic-4.tptp
fof(a, axiom, p).
fof(goal, conjecture, p).
\end{tptp}
    \item $p ∧ q ⊢ q ∧ p$
\begin{tptp}
$ cat conj-3.tptp
fof(a, axiom, p & q).
fof(goal, conjecture, q & p).
\end{tptp}
    \item $⊢ ¬ (p ∧ ¬ p) ∨ (q ∧ ¬ q)$
\begin{tptp}
$ cat neg-7.tptp
fof(goal, conjecture, ~ ((p & ~ p) | (q & ~ q))).
\end{tptp}
    \end{itemize}
\end{frame}

@@

\begin{frame}[fragile]{TSTP Syntax}
\label{tstp-syntax}
 \tikz[overlay,remember picture]
  \node at (0.92\textwidth, -0.5cm)
    {\includegraphics[width=0.15\textwidth]{figures/tstp}};

A TSTP derivation
\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/QuickGuide/Derivations.html}}
\begin{itemize}
  \item Is a Directed Acyclic Graph where
  \begin{itemize}
    \item[\texttt{leaf}] is a formulae from the TPTP input
    \item[\texttt{node}] is a formulae inferred from parent formulae
    \item[\texttt{root}] the final derived formulae
  \end{itemize}
  \item Is a list of annotated formulae:
  {\footnotesize
    \begin{center}
\begin{verbatim}
language(name,role,formula,source[,useful info]).
\end{verbatim}
  \end{center}}
\begin{itemize}
  \item[\texttt{source}] typically is a inference record:
    {\footnotesize
\begin{verbatim}
  inference(rule, [], [])
  \end{verbatim}}
\begin{itemize}
    \item[-] The inference rule name,
    \item[-] a list of useful inference information, and
    \item[-] a list of references to its parent formulae.
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{TSTP Example}
\label{tstp-example}
 \tikz[overlay,remember picture]
  \node at (0.93\textwidth, -0.3cm)
    {\includegraphics[width=0.15\textwidth]{figures/metis}};

@@

\begin{frame}[fragile]{DAG for the previous TSTP derivation found by \texttt{Metis} ATP}
\label{tstp-dag}
\vfill
\begin{center}
\includegraphics[height=0.8\textheight]{figures/derivation.png}
\end{center}
\end{frame}

\begin{frame}{Design decisions for the Reconstruction Tool}
\label{agda}

 \tikz[overlay,remember picture]
  \node at (0.93\textwidth, 0cm)
    {\includegraphics[width=0.2\textwidth]{figures/athena}};

\textbf{Programming Languages}
\begin{itemize}
\item \texttt{Haskell} is a standardized, general-purpose purely functional programming language.\\
Our usages:
\begin{itemize}
    \item Parsing
    \item AST construction
    \item Creation and analysis of DAG derivations
    \item Analysis of inference rules used
    \item Generation of Agda code of the proof
\end{itemize}

\item \texttt{Agda} is dependently typed functional programming language and it also a proof assistant.\\
Our usages:
    \begin{itemize}
    \item Logic framework for Classical Propositional Logic
    \item Type-Checker for the Agda code of the proofs generated
    \end{itemize}
\end{itemize}

\begin{frame}{References}
\label{references}
\printbibliography
\end{frame}
